---
title: "Practical Machine Learning Course Project"
author: "Vladimir"
date: "21 Dec 2014"
output: html_document
---

Online version available here http://rpubs.com/vladlet/51405

## Introduction 

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement—a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data 
```{r results='hide'}
library(caret)
library(randomForest)
````

Download data
```{r}
trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
````

```{r}
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="curl")
}
````
Read the files
```{r}
data <- read.csv(trainFile, na.string=c("NA", ""))
check <- read.csv(testFile, na.string=c("NA", ""))
````

Clean data 
```{r}
classe <- data$classe
data <- data[,names(data[,colSums(is.na(data))==0])]
data <- data[, sapply(data, is.numeric)]
remove <- grepl("^X|timestamp", names(data))
data <- data[, !remove]
pr_is <- check$problem_id
check <- check[, names(data)]
check$problem_id <- pr_is
data$classe <- classe
````

Init random seed
```{r}
set.seed(19634)
````
Split data to training and testing for cross validation.
```{r}
inTrain <- createDataPartition(data$classe, p = 0.75, list = F)
training <- data[inTrain,]
testing <- data[-inTrain,]
````

## Model building
Fit a model using random forest
```{r, cache=TRUE}
ctrlRf <- trainControl(method = "oob")
modelRf <- train(classe ~ ., training, method = "rf", ntree = 200, trControl = ctrlRf)
````
## Cross Validation Testing and Out-of-Sample Error Estimate
```{r}
fitRf <- predict(modelRf, testing)
confusionMatrix(testing$classe, fitRf)
```
```{r}
accur <- postResample(fitRf, testing$classe)
accur
```
The estimated accuracy of the model is `r accur[[1]]` and the estimated out-of-sample error based on our fitted model applied to the cross validation dataset is `r 1-accur[[1]]`.

## Predicted Results
Finally, we apply the pre-processing to the original testing dataset, after removing the extraneous column labeled `problem_id` 

```{r}
find <- which(names(check) == "problem_id")
finalRf = predict(modelRf, check[, -find])
finalRf
```